Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-04-03 conclusion.tex
klunky first pass at conclusions.

>>>>

HEP was amongst the first, if not the very first experimental community to
realize the importance of using WMS to manage their computational campaign(s).
As computing become increasingly critical for a range of other experiments,
the impact of WMS on HEP foreshadows the importance of WMS for other
experiments and their computational campaigns. Other experiments such as SKA,
LSST etc.., will have their own workload characteristics, resources types and
federation constraints, as well metrics of performance.  The design and
experience captured in this paper should proof invaluable for computational
campaigns that need to evaluate the relative merits of different approaches.

Thanks to the efforts captured in this paper, the time-averaged utilization of
Titan has consistently been upwards of 90\%. Compare this to the time averaged
utilization of other similar leadership machines such as NSF's flagship Blue
Waters, where the average utilization fluctuates between 60-80\% (XDMoD)
machine~\cite{bw-sucks}. These impressive if not unprecedented operation gains
aside, this work is just a starting point towards more effective operational
models for future leadership and online analytical platforms\cite{foap-url},
which will have to support ever increasing complex workloads with varying
models for dynamic resource federation. Recent efforts with Google Cloud for
the CMS experiment -- HEPCloud~\cite{hepcloud,google-hep} represent similar
yet different approaches to dynamic resource federation.

Our experience demonstrated the advantages arising from a  systems engineering
and abstractions based approach, in particular the pilot abstraction. As the
type, heterogeneity and complexity of future platforms  (e.g., Exascale,
Commercial Clouds) for science increases, there will be a greater need and
emphasis on abstractions to ensure WMS can both utilize as well as support
platforms.

### 2017-04-03 introduction.tex
klunky first pass at conclusions.

This paper is a case study of how the ATLAS experiment has "broken free" of
the traditional computational approach of high-throughput computing on
distributed resources to embrace new platforms, in particular high-performance
computers. Specifically, we discuss the experience of integrating the PanDA
workload management system with Titan ~\textemdash{} a DOE leadership
computing facility. Consequently, Titan now analyzes up to 5\% of workload  of
the ATLAS experiment. The case study presents the design and integration of
PanDA to reach sustained production scales (ca. 100 M core-hours a years). It
also discusses the investigation of a task execution runtime system on Titan,
that is based on  the pilot-abstractions and which allows advanced execution
modes of the ATLAS workload as well as the enhanced support for heterogeneous
workloads (such as molecular dynamics).

This state-of-practice paper provides multiple contributions.  It (i)
documents the many design and operational considerations that have been  taken
to support the sustained, scalable and production usage of Titan for
historically high-throughput workloads, (ii) Extensions to PanDA to support
non-traditional heterogeneous workloads and execution modes,  and (iii) As the
community looks forward to designing the next generation of online analytical
platforms~\cite{foap}, this project provides some early lessons and guidance
for how current and future experimental and observational systems can be
integrated with production supercomputers in a general and extensible manner.

>>>>

This paper is a case study of how the ATLAS experiment has "broken free" of
the traditional computational approach of high-throughput computing on
distributed resources to embrace new platforms, in particular high-performance
computers. Specifically, we discuss the experience of integrating the PanDA
workload management system with Titan ~\textemdash{} a DOE leadership
computing facility. Consequently, Titan now analyzes up to 5\% of the
simulation workload  of the ATLAS experiment. The case study presents the
design and integration of PanDA to reach sustained production scales (ca. 100
M core-hours a years). It also discusses the investigation of a task execution
runtime system on Titan, that is based on  the pilot-abstractions and which
allows advanced execution modes of the ATLAS workload as well as the enhanced
support for heterogeneous workloads (such as molecular dynamics).

This state-of-practice paper provides multiple contributions.  It (i)
documents the many design and operational considerations that have been  taken
to support the sustained, scalable and production usage of Titan for
historically high-throughput workloads, (ii) Extensions to PanDA to support
non-traditional heterogeneous workloads and execution modes,  and (iii) As the
community looks forward to designing the next generation of online analytical
platforms~\cite{foap-url}, this project provides some early lessons and guidance
for how current and future experimental and observational systems can be
integrated with production supercomputers in a general and extensible manner.
